---
title: "4: Timing benchmarks"
author: "Martin R. Smith <martin.smith@durham.ac.uk>"
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    toc: no
  rmarkdown::html_vignette:
    default: yes
bibliography: ../inst/REFERENCES.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-old-doi-prefix.csl
---

```{r pc-specs, echo = FALSE}
if (requireNamespace('benchmarkme', quietly = TRUE)) {
  cpu <- benchmarkme::get_cpu()
  ram <- benchmarkme::get_ram()
} else {
  cpu <- list(model_name = "UNKNOWN", no_of_cores = "UNKNOWN")
  ram <- "UNKNOWN"
}
```


```{r setup, echo = FALSE, message = FALSE}
library('TreeDist')
path.dist <- phangorn::path.dist
SPR.dist <- phangorn::SPR.dist
TBRDist <- TBRDist::TBRDist
RF <- TreeDist::RobinsonFoulds

nTrees <- 45L # For c. 1000 comparisons
nRepeats <- 1L # Increase locally for slower but more accurate benchmarking
nPairs <- nTrees * (nTrees - 1) / 2

SPRWalk <- function (nTip) {
  set.seed(0L)
  sprWalk <- vector('list', nTrees)
  sprWalk[[1]] <- lastTree <- TreeTools::PectinateTree(nTip)
  
  for (i in seq_len(nTrees)[-1]) {
    sprWalk[[i]] <- lastTree <- TreeSearch::SPR(lastTree)
  }
  
  trees <- lapply(sprWalk, TreeTools::Preorder)
  trees <- lapply(trees, TreeTools::Cladewise)
  class(trees) <- 'multiPhylo'
  trees
}

Times <- function (trees) {
  splits <- TreeTools::as.Splits(trees)
  timings <- microbenchmark::microbenchmark(
    dpi = SharedPhylogeneticInfo(splits, normalize = FALSE),
    msid = MatchingSplitInfo(splits, normalize = FALSE),
    cid = MutualClusteringInfo(splits, normalize = FALSE),
    nea = NyeTreeSimilarity(splits, normalize = FALSE),
    ja2 = JaccardRobinsonFoulds(splits, k=2L, similarity = TRUE, normalize = FALSE),
    ja4 = JaccardRobinsonFoulds(splits, k=4L, similarity = TRUE, normalize = FALSE),
    jna2 = JaccardRobinsonFoulds(splits, k=2L, similarity = TRUE, arboreal = FALSE, normalize = FALSE),
    jna4 = JaccardRobinsonFoulds(splits, k=4L, similarity = TRUE, arboreal = FALSE, normalize = FALSE),
    msd = MatchingSplitDistance(splits),
    qd = Quartet::ManyToManyQuartetAgreement(trees),
    mast = TreeDistData::PairwiseDistances(trees, MASTSize, rooted = FALSE),
    nni = TreeDistData::PairwiseDistances(trees, NNIDist, 3L),
    spr = SPR.dist(trees),
    tbr = TBRDist(trees, exact = FALSE),
    rf = RF(splits),
    rfi = InfoRobinsonFoulds(splits),
    path = path.dist(trees),
    
    times = nRepeats
  )
}

Plot <- function (timings, nTip) {
  par(cex = 0.8)
  boxplot(timings, border = TreeDistData::TreeDistCol(
    as.character(summary(timings)[, 'expr'])),
        xlab = 'Method', ylab = 'Total time elapsed / ms',
        main = paste(nTip, 'leaves'))
}

```

The time taken to compare all `r nPairs` pairs of trees by each method, 
was calculated, replicating timing `r nRepeats` times.

The below are the results of a benchmarking exercise conducted on an
`r cpu$model_name` machine with
<!--`r cpu$no_of_cores` cores and-->
`r as.character(benchmarkme:::print.ram(ram * 1024))` of RAM,
comparing `r nTrees` trees obtained by performing successive subtree pruning and
regrafting rearrangements on a pectinate starting tree.  
The values reported in Smith [-@Smith2020] were calculated on a modest desktop computer,
and will differ from those reported here, which will have been generated on
the hardware system used to render the documentation.

## Figure

```{r timings-20, echo = FALSE, output = 'asis', fig.width = 8, fig.height = 4}
times20 <- Times(SPRWalk(20))
Plot(times20, 20)
```

```{r timings-50, echo = FALSE, output = 'asis', fig.width = 8, fig.height = 4}
times50 <- Times(SPRWalk(50))
Plot(times50, 50)
```

## Tabulation

```{r timings-table, echo = FALSE}
results <- summary(times20, unit = 'us')
#cols <- c('mean', 'median')
cols <- c('mean')
rownames(results) <- TreeDistData::tdAbbrevs[as.character(results[, 'expr'])]
results <- cbind(results[, cols, drop = FALSE],
                 summary(times50, unit = 'us')[, cols, drop = FALSE])

colnames (results) <- paste(rep(c(20, 50), each = length(cols)), 'leaves,', cols)
knitr::kable(signif(results[order(results[, 1]), ] / nPairs, 2),
             caption = "Time per comparison / µs")
```

## Reference
