---
title: "Supp. Figs: Metric correlation"
author: "Martin R. Smith <martin.smith@durham.ac.uk>"
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    toc: no
  rmarkdown::html_vignette:
    default: yes
---

# Correlation between metrics

```{r data-load, echo=FALSE}
library('TreeDistData')
data("distanceDistribution25", package="TreeDistData")
# No point in plotting both MAST and MASTI, as they're perfect rank correlates
exclude <- c('nni_t', 'nni_u', 'nni_l', 'masti')
distanceDistribution25 <- distanceDistribution25[!rownames(distanceDistribution25) %in% exclude, ]
data("distanceDistribution50", package="TreeDistData")
distanceDistribution50 <- distanceDistribution50[!rownames(distanceDistribution50) %in% exclude, ]
toCorrelate <- c(
                 "path",
                 "qd", 
                 "msd", "msid", 
                 "dpi", "cid", "nts",
                 "ja2", "ja4", "jna2", "jna4",
                 "rfi", "rf",
                 "spr", "tbr_l", "tbr_u", "mast" # Integers
                 )
if (!all(rownames(distanceDistribution50) %in% toCorrelate)) {
  warning("Missing ", rownames(distanceDistribution50)[!rownames(distanceDistribution50) %in% toCorrelate])
}
colScale <- colorspace::sequential_hcl(1000, palette='viridis')
```

```{r functions, echo=FALSE}
CorrPlot <- function (distribution, methodA, methodB) {
  a <- distribution[methodA, ]
  b <- distribution[methodB, ]
  plot(a ~ b, pch='.', col='#00000033', axes=FALSE, ylab='', xlab='')
  reg <- lm(a ~ b)
  abline(reg, col='#ff3333')
}

PlotR2 <- function (distribution, methodA, methodB) {
  a <- distribution[methodA, ]
  b <- distribution[methodB, ]
  plot.new()
  reg <- lm(a ~ b)
  r2 <- summary(reg)$adj.r.squared
  regCol <- colScale[max(ceiling((r2 - 0.5) * 2000), 1L)]
  text(0.5, 0.5, format(r2, digits = 2, nsmall = 2), col = regCol, cex = 1.4)
}

PlotCorr <- function(distribution) {
  par(mfrow = length(toCorrelate) + c(1L, 0L), mar=rep(0.2, 4))
  lapply(seq_along(toCorrelate), function(i)
    lapply(seq_along(toCorrelate), function(j) {
      if (i == j) {
        plot.new()
        method <- toCorrelate[i]
        text(0.5, 0.5, tdBoxAbbrevs[method],
             col = TreeDistCol(method), cex = 0.9)
        box(which = 'plot', col = TreeDistCol(method), lwd = 3)
      } else if (i < j) {
        PlotR2(distribution, toCorrelate[i], toCorrelate[j])
      } else {
        CorrPlot(distribution, toCorrelate[i], toCorrelate[j])
      }
    })) -> XX
}
```

```{R metric-correlation, echo=FALSE, fig.width=7, fig.asp=9/8, out.width='98%', fig.align='center', warning=FALSE}
PlotCorr(distanceDistribution25)
```

**Supplementary Figure 1. Correlation (adjusted r²) between distances for 10&nbsp;000 pairs of 25-leaf trees**

```{R metric-correlation-50, echo=FALSE, fig.width=7, fig.asp=9/8, out.width='98%', fig.align='center', warning=FALSE}
PlotCorr(distanceDistribution50)
```

**Supplementary Figure 2. Correlation (adjusted r²) between distances for 10&nbsp;000 pairs of 50-leaf trees**


Not plotted:

- MASTI, which correlates perfectly with MAST;

- NNI: both upper and lower bounds correlate almost perfectly with RF, due to
  the method of calculating the approximation.

