---
title: "SPR walking"
author: "Martin R. Smith <martin.smith@durham.ac.uk>"
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    toc: no
  rmarkdown::html_vignette:
    default: yes
bibliography: ../inst/REFERENCES.bib
csl: ../inst/apa-old-doi-prefix.csl
vignette: >
  %\VignetteIndexEntry{SPR walking}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r load, echo = FALSE, message=FALSE}
# Update from data-raw/sprDistances.R
nTrees <- 100L
nTip <- 50L

moves <- abs(outer(seq_len(nTrees), seq_len(nTrees), '-'))
distances <- TreeDistData::sprDistances
plotSequence <- tdMethods[tdMethods %in% names(distances)]

Plotter <- function (method) {
  dat <- as.matrix(distances[[method]])
  if (method %in% c('mast', 'masti')) dat <- nTip - dat
  dat.A <- dat[!is.na(dat)]
  moves.A <- moves[!is.na(dat)]
  col <- TreeDistCol(method)
  boxplot(dat.A ~ moves.A, border = col, 
          ylab = TreeDistData::tdAbbrevs[method],
          xlab = 'SPR moves', axes= F)
  tau <- cor.test(dat.A, moves.A, method="kendall", use="pairwise")$estimate
  text(max(moves.A), min(dat.A), paste('tau-b =', signif(tau, 3)), pos = 2, col = col)
  legend('topleft', TreeDistData::tdAbbrevs[method], bty='n', col = col)
}
```

SPR Walking: generating `r nTrees` trees with `r nTip` tips.
Start with a pectinate `r nTip`-leaf tree.  
Each subsequent tree is one SPR rearrangement away from the previous one.
Trees SPR rearrangements continue until a forest of `r nTrees` trees has been
generated.  Distances between each pair of trees are then calculated, and 
plotted against the number of SPR moves that separate the pair.

```{r plots, echo = FALSE, fig.width=7, fig.height=6}
par(mfrow = c(3, 7), mar = c(1, 0.1, 1, 0.1), oma=c(0, 0, 0, 0))
XX <- lapply(plotSequence, Plotter)
```

We then calculate the rank correlation between the measured tree distance and 
the number of SPR moves used to generate the trees – which may not correspond 
to the exact SPR distance, as one SPR move may 'undo' another.

Higher coefficients of correlation are, to a first approximation, desirable.

```{r table, output="asis", echo = FALSE}
Tabler <- function (method) {
  dat <- as.matrix(distances[[method]])
  cor.test(dat[!is.na(dat)], moves[!is.na(dat)], method="kendall",
           use="pairwise")$estimate
}

knitr::kable(t(abs(signif(vapply(plotSequence, Tabler, double(1)), 3))))
```
